---
title: "GGplot"
output: pdf_document
---

## Simulate some data

## 3 Factor Variables
FacVar1=as.factor(rep(c("level1","level2"),25))
FacVar2=as.factor(rep(c("levelA","levelB","levelC"),17)[-51])
FacVar3=as.factor(rep(c("levelI","levelII","levelIII","levelIV"),13)[-c(51:52)])

## 4 Numeric Vars
set.seed(123)
NumVar1=round(rnorm(n=50,mean=1000,sd=50),digits=2) ## Normal distribution
set.seed(123)
NumVar2=round(runif(n=50,min=500,max=1500),digits=2) ## Uniform distribution
set.seed(123)
NumVar3=round(rexp(n=50,rate=.001)) ## Exponential distribution
NumVar4=2001:2050

simData=data.frame(FacVar1,FacVar2,FacVar3,NumVar1,NumVar2,NumVar3,NumVar4)



library(ggplot2)
library(reshape2)

#What Is GGplot2
Ggplot2 is a plotting system for R, based on the grammar of graphics, which tries to take the good parts of base and lattice graphics and none of the bad parts. It takes care of many of the fiddly details that make plotting a hassle (like drawing legends) as well as providing a powerful model of graphics that makes it easy to produce complex multi-layered graphics

#Graphing One Variable
##Numberic
###Index Plot
```{r}
ggplot(simData,aes(y=NumVar1,x=1:nrow(simData),group="NumVar1"))+geom_point()+geom_line()+ xlab("") ## Index plot
```


###Histogram
```{r}
ggplot(simData,aes(x=NumVar1))+geom_histogram() 
```


###Kernel Density Plot
```{r}
ggplot(simData,aes(x=NumVar1))+geom_density() 
```


###Box Plot
```{r}
ggplot(simData,aes(x=factor(""),y=NumVar1))+geom_boxplot()+ xlab("") ## box plot
```


##Factor

### barplot
```{r}
ggplot(simData,aes(x=FacVar3))+geom_bar()
```

###Pie Chart
```{r}
ggplot(simData,aes(x = factor(""), fill=FacVar3, label=FacVar3))+geom_bar()+ coord_polar(theta = "y")  +scale_x_discrete("")
```
#Graphing Two Variables
## Two Numeric

```{r, echo=FALSE}
simtmp=simData[,c(4:5)] ## 4th and 5th columns are NumVar1 and NumVar2
simtmp$index=1:nrow(simtmp)
simtmpmelt=melt(simtmp,id=c("index"))

### line plots with observation number as index
ggplot(simtmpmelt,aes(y=value,x=index,color=variable))+geom_point()+geom_line()+xlab("")
```

###Density Plot
```{r}
ggplot(simtmpmelt,aes(x=value,color=variable))+geom_density()
```

###Scatter Plot
```{r}
ggplot(simData,aes(x=NumVar1,y=NumVar2))+geom_point()
```

##Factor

###Mosaic plot
```{r}
ggMMplot <- function(var1, var2){
  require(ggplot2)
  levVar1 <- length(levels(var1))
  levVar2 <- length(levels(var2))
  
  jointTable <- prop.table(table(var1, var2))
  plotData <- as.data.frame(jointTable)
  plotData$marginVar1 <- prop.table(table(var1))
  plotData$var2Height <- plotData$Freq / plotData$marginVar1
  plotData$var1Center <- c(0, cumsum(plotData$marginVar1)[1:levVar1 -1]) +
    plotData$marginVar1 / 2
  
  ggplot(plotData, aes(var1Center, var2Height)) +
    geom_bar(stat = "identity", aes(width = marginVar1, fill = var2), col = "Black") +
    geom_text(aes(label = as.character(var1), x = var1Center, y = 1.05)) 
}
ggMMplot(simData$FacVar2, simData$FacVar3)
```

###Barplot
####To Get the Cross Tab Enter:
bartabledat = as.data.frame(table(simData$FacVar2, simData$FacVar3)) 
####To Plot:
ggplot(bartabledat,aes(x=Var2,y=Freq,fill=Var1))+geom_bar(position="dodge") 


```{r, echo=FALSE}
bartabledat = as.data.frame(table(simData$FacVar2, simData$FacVar3)) ## get the cross tab
ggplot(bartabledat,aes(x=Var2,y=Freq,fill=Var1))+geom_bar(position="dodge") ## plot
```


###Stacked Barplot
ggplot(bartabledat,aes(x=Var2,y=Freq,fill=Var1))+geom_bar()

###100% Stacked Barplot
bartableprop =as.data.frame(prop.table(table(simData$FacVar2, simData$FacVar3),2)*100) 
ggplot(bartableprop,aes(x=Var2,y=Freq,fill=Var1))+geom_bar()

##One Numberic One Factor
